---
title: "2 - Vectors in R"
author: "Kamal Chouhbi"
output:
   html_notebook:
    theme: simplex
    toc: true
    toc_depth: 4
---

A **vector** is a data type that is used to store a collection of elements. The elements in a vector can be of any basic type, but all elements of a single vector must be of the same type. In this class, we will primarily work with vectors of the **numeric** data type. 

### **Creating Vectors**

There are several ways to create a vector in R. In this notebook we will discuss two functions that can be used to create vectors from scratch: The `c` function, and the 

#### **Combine**

Vectors can be formed using the `c()` function. The name of this function is short for the word "combine". In the code chunk below, we will create two vectors name `v1` and `v2`, each of which contains four elements. 


```{r}
v1 <- c(3, 1, 4, 1)
v2 <- c(5, 9, 2, 6)
```

We will view the contents of the vector `v1`.  

```{r}
v1
```

Similarly, we can view the contents of 'v2'. 

```{r}
v2
```

We can use the `c()` function to combine two vectors into a larger vector.

```{r}
v3 <- c(v1, v2)
v3
```

Although we will mostly work with numeric vectors, we will occasionally need to work with vectors whose elements are of the character class. 

```{r}
text_vec <- c("Alex", "Beth", "Craig", "Drew", "Emma")
text_vec
```

#### **The `seq` Function**

The `seq()` function is a convenient tool for creating a vector consisting of several values, evently spaced over a range. 

```{r}
s1 <- seq(from=0, to=5, by=0.2)
s1
```

```{r}
s2 <- seq(from=6, to=10, by=0.5)
s2
```

If the terms of our sequence vary by 1, we can use the `start:end` shorthand notation to create the sequence. 

```{r}
s3 <- 3:12
s3
```


#### **Functions**

In this section, we will introduce several functions that can be used to get information about a vector. The first such function that we will demostrate is `length()`, which returns the number of elements in the vector. 

```{r}
length(s1)
```

The `sum()` function can be used to sum up the elements of a numeric vector. 

```{r}
sum(s1)
```

Being a statistical programming language, R also provides several functions for calculating statistics for the numbers stored in a vector. For example, the `mean()` function returns the arithmetic average of the elements of a vector. 

```{r}
mean(s1)
```

We can use `sd()` to find the sample standard deviation of the numbers in a vector. 

```{r}
sd(s1)
```

And `var()` provides the sample variance.

```{r}
var(s1)
```

### **Vector Operations**

One of the biggest advantages of vectors is that they support elementwise operations. If we wish to apply a specific operation to each element in a vector, we can simply apply the operation to the vector as a whole. R will automatically apply the operation to each element of the vector, without the need for using a loop. 

The follow cell illustrates the operation of scalar multiplication. We will multiply the contents of our vector `v1` (defined at the beginning of the notebook) by the scalar value 4. 

```{r}
4 * v1
```

We can also add a constant amount to all of the entries in a vector. 

```{r}
30 + v1
```


If we add two vectors of the same length, then the addition will be performed elementwise. 

```{r}
v1 + v2
```

We can also perform elementwise multiplication, division, or subtraction. The following cell demonstrates elementwise multiplication.

```{r}
v1 * v2
```

We can also use the `^` operator to raise every element of a vector to a common power. 

```{r}
v1^2
```


#### **Example: Calculating Revenue**

Assume that an online retailer sells four products: Product A, Product B, Product C, and Product D. In the cell below, we define two vectors: `units` and `price`. Assume that the vector `units` stores the number of units that the retailer sold of Products A - D during a given week, listed in alphabetical order. The vector `price` lists the price of the four products, in the same order. The following cell calculates the total revenue for the retailer during the week. 

```{r}
units <- c(217, 362, 178, 255)
price <- c(19.99, 15.49, 12.79, 22.49)
revenue <- sum(units * price)
revenue
```

#### **Example: Projectile Motion**

Assume that a ball is launched upward from a height of 640 feet with an initial velocity of 96 ft/sec. Then the height `h` of the ball after `t` seconds is given by the formula: 

$$h(t) = 640 + 96 t - 16 t^2$$
The following code chunk calcualtes the height of the ball at times $t = 0, 1, 2, ..., 10$. 


```{r}
t <- 0:10
h <- 640 + 96*t - 16*t^2
h
```

It is occasionally useful to show the contents of two vectors of the same length stacked together horizontally. The function `rbind()` can be used to accomplish this. 

```{r}
rbind(t, h)
```

#### **Example: SSE and MSE**

Two vector calculations that are of particular importance in statistics and machine learning are to find the sum and mean of the squares of the entries in a vector. If the elements of the vector represent errors associated with several individual predictions, the we called the resulting values the **Sum of Square Errors (SSE)** and **Mean Squared Error (MSE)**. 

In the cell below, we define an error vector `e`.

```{r}
e <- c(1.7, 2.1, -3.2, 4.1, -2.7)
```

In the next cell, we will calculate `SSE`. 

```{r}
SSE <- sum(e^2)
SSE
```

The following cell calculates the value of `MSE`. 

```{r}
MSE <- mean(e^2)
MSE
```
