---
title: "4 - Random Sampling in R"
author: "Kamal Chouhbi"
output:
  html_notebook:
    theme: simplex
    toc: true
    toc_depth: 4
---

There are often occasions when it is useful to take a random sample from an already existing vector, or to generate data according to some distribution. We will discuss methods for creating randomly generated vectors in this notebook.

### **Sampling from a Vector**

We can use the `sample()` function to randomly select some number of elements from a vector. We can control the number of elements selected using the `size` parameter. 

```{r}
v <- c(71, 58, 38, 99, 48, 55,
       25, 67, 32, 36, 84, 18)

sample(v, size=3)
```

Note that we will get a different result each time we run the `sample` command. 

```{r}
sample(v, size=3)
```

By default, the sampling is performed without replacement. We can alter this behavior by setting `replace=TRUE`. 

```{r}
sample(v, size=6, replace=TRUE)
```


### **Sampling from a Normal Distribution**

We can use the `rnorm` function to generate a random sample according to a normal distribution. In the cell below, we generate 300 values from a normal distribution with a mean of 20 and a standard deviation of 4. We then display the first 10 values in our random sample. 

```{r}
sample_1 <- rnorm(n=300, mean=20, sd=4)
sample_1[1:10]
```

It is difficult to tell that our sample is randomly generate just by looking at the values for 10 (or all 300) observations. However, recall that for a normal distribution:

* 68.26% of all observations are within 1 standard deviation from the mean.
* 95.44% of all observations are within 1 standard deviation from the mean. 
* 99.73% of all observations are within 1 standard deviation from the mean. 

Since we are working with a randomly generated sample, our proportions might vary a bit from this, but they should be similar. 

```{r}
mean((sample_1 > 16) & (sample_1 < 24))
```

```{r}
mean((sample_1 > 12) & (sample_1 < 28))
```

```{r}
mean((sample_1 > 8) & (sample_1 < 32))
```

### **Sampling from a Uniform Distribution**

We can use the function `runif` to sample from a uniform distribution. In the cell below, we generate 300 observations of a variable that is uniformly distributed on the interval `[0, 10]`. We then display the first 10 values in our random sample. 

```{r}
sample_2 <- runif(n=300, min=0, max=10)
sample_2[1:10]
```

Since the length of our interval is 10, roughly 70% of our observations should be less than 7. 

```{r}
mean(sample_2 < 7)
```

Roughly 50% of our observations should be between 3 and 8. 

```{r}
mean((sample_2 > 3) & (sample_2 < 8))
```

Roughly 80% of our observations should be greater than 2. 

```{r}
mean(sample_2 > 2)
```